---
title: "BildAktVR data analysis 1"
author: "Ondrej Havlicek"
output:
  html_notebook:
    code_folding: hide
    df_print: paged
    theme: flatly
    toc: yes
---
[ohavlicek@gmail.com](mailto:ohavlicek@gmail.com) 11/2017

## Preparations
### Load packages
```{r Libraries, message=FALSE, warning=FALSE}
library(tidyverse)  #all basic functions
library(forcats) #working with factors
library(lme4) #mixed effects models
library(lmerTest) #mixed effects models pvalues
library(sjPlot) #liner models tables
```

### Defititions
```{r Definitions}
datapath <- file.path(getwd(), "Data", "Main")
datafilepattern <- "BildAkt.*\\.csv"

#ggplot2 theme
#old_theme <- theme_set(theme_grey())
```

### Read data and convert column types
```{r Read_data, message=FALSE, warning=FALSE}
datafiles <- list.files(datapath, datafilepattern, full.names = T)

rawdata <- do.call(rbind, lapply(datafiles, read_delim, delim=";")) # read all files and combine them into one tibble

rawdata <- rawdata %>% mutate(firstSceneType = parse_factor(firstSceneType, c("3D", "2D")),
                   firstEventType = parse_factor(firstEventType, c("Action", "Control")),
                   isPracticeBlock = as.logical(isPracticeBlock),
                   sceneType = parse_factor(setupType, c("3D", "2D")),
                   eventType = parse_factor(eventType, c("Action", "Control")))
```

### Prepare subset of data for analysis, create new columns
```{r Subset}
data <- rawdata %>% filter(isPracticeBlock == FALSE) %>% 
  select(Subject = subject,
         Session = session,
         FirstSceneType = firstSceneType,
         FirstEventType = firstEventType,
         Block = blockNo,
         SceneType = sceneType,
         EventType = eventType,
         Trial = trialNoInBlock,
         TrialTotal = trialNoTotal,
         Objects = noObjects,
         ObjectsResp = responseObjectsNo,
         Interval = interval,
         IntervalResp = responseInterval,
         # TODO: add TSTrialOnset
         TSObjectsResp = responseObjectsTimestamp,
         TSEvent = eventTimestamp,
         TSTone = toneTimestamp, ##TSTone is recorded when sending the command playDelayed to produce the tone after a specified interval, so tone in fact occurs later, at an uknown time, but roughly TSTone + Interval
         TSIntervalResp = responseIntervalTimestamp
  ) %>% mutate(#TODO: ObjectsRT = TSObjectsResp - TSTrialOnset,
    ObjectsError = ObjectsResp - Objects,
    IntervalError = IntervalResp - Interval,
    IntervalRT = TSIntervalResp - (TSTone), #correct? apparently, no need to adjust the RT by Interval?? it just gets regressed out...
    #IntervalRT2 = TSIntervalResp - (TSEvent + Interval)
    IntervalLogRT = log(IntervalRT)
    ) 

data #preview
```

## Exploratory analysis
### Explore a bit scenes distribution and randomization
```{r Scene_interval_randomization}
#with(subset(rawdata, isPracticeBlock==F), table(sceneId, blockNo, subject)) #is each scene used only once within a block? seems so..
#with(subset(rawdata, isPracticeBlock==F), table(sceneId, interval, subject)) #is interval assigned to scenes randomly? usually just two different intervals per scene..
#with(subset(rawdata, isPracticeBlock==F), table(sceneId, subject)) #how often is a scene used per subject?
#are different intervals equally distributed with object numbers?
data %>% ggplot(., aes(x=Objects, y=Interval)) + geom_bin2d() + scale_y_continuous(breaks=c(150, 350, 550, 750)) + scale_x_continuous(breaks=6:15) +
  scale_fill_gradientn(colours=rainbow(4))
#with(data, table(Objects, Interval)) #seems random, although far from uniformly random
#with(data, cor.test(Objects, Interval)) #tiny negative correlation
```
Number of objects inside a scene is not orthogonal to time interval: some combinations appear often, some less often. We will have to revise the VR code.


### Objects error exploration
Distributions of object estimation errors (error = estimated - actual) for each participant:
```{r Objects_dists}
data %>% ggplot(., aes(x=ObjectsError, linetype=SceneType, color=EventType)) +  
  facet_wrap(~Subject) + geom_freqpoly(bins=5) #geom_density(alpha=0.3)
```

People are mostly correct about the object count.

```{r Objects_accuracy}
objErrors <- data %>% group_by(Subject, SceneType, EventType) %>% 
  summarize(NTrials = n(), ObjectsErrorMean = mean(ObjectsError), ObjectsErrorMedian = median(ObjectsError), ObjectsErrorAbsMean = mean(abs(ObjectsError)), ObjectsErrorSD = sd(ObjectsError), ObjectsErrorCor = cor(Objects,ObjectsResp)) 

# Accuracy
objErrors %>%  ggplot(., aes(y=ObjectsErrorMean, x=SceneType)) +
  geom_boxplot() + facet_grid(.~EventType)
```

Numbers of objects are being underestimated in 2D scenes: This could be a confound for interpretation of any interval estimation effect - 3D and 2D scenes differ in object counting task difficulty. Should we reduce overlap of objects in the scenes?

```{r Objects_precision}
# Precision
objErrors %>%  ggplot(., aes(y=ObjectsErrorSD, x=SceneType)) +
  geom_boxplot() + facet_grid(.~EventType)
```

Precision is also better for 3D than for 2D scenes

```{r Objects_correlation}
# Correlation, discrimination
objErrors %>%  ggplot(., aes(y=ObjectsErrorCor, x=SceneType)) +
  geom_boxplot() + facet_grid(.~EventType)
```

There is also a difference in correlation between actual and estimated number of objects, but very tiny.


### Objects mixed-effects model
ObjectsError ~ Objects + SceneType*EventType + (1+Objects + SceneType + EventType|Subject)
```{r Objects_mixed-effects_model, message=FALSE, warning=FALSE}
objectsLmer <- as.data.frame(data) %>% lmer(ObjectsError ~ Objects + SceneType * EventType + (1+Objects + SceneType + EventType|Subject), .)
#random interaction fails to converge
#summary(objectsLmer)

sjt.lmer(objectsLmer, file="objectsLmer.html")# We have to save the table in html format.
htmltools::includeHTML("objectsLmer.html") #And include it later, not needed for html_document type, only for notebook
```
The error in object count is not (sig.) influenced by number of objects in the scene, but is influenced by the scene type (2D = more underestimation of the count)

### Intervals error exploration
```{r Intervals_dens}
# Are interval errors approx normally distributed?
data %>% ggplot(., aes(x=IntervalError, linetype=SceneType, color=EventType)) +  
  facet_wrap(~Subject) + geom_freqpoly(bins=15) #geom_density(alpha=0.3)
```

Interval estimation errors (Estimated - Actual) are pretty much normally distributed with pretty much similar variance across subjects and conditions

```{r Intervals_accuracy}
intErrors <- data %>% group_by(Subject, SceneType, EventType) %>% 
  summarize(NTrials = n(), IntervalErrorMean = mean(IntervalError), IntervalErrorMedian = median(IntervalError), IntervalErrorAbsMean = mean(abs(IntervalError)), IntervalErrorSD = sd(IntervalError), IntervalErrorCor = cor(Interval,IntervalResp)) 

# Accuracy
intErrors %>%  ggplot(., aes(y=IntervalErrorMean, x=SceneType)) +
  geom_boxplot() + facet_grid(.~EventType)
```

Intervals are judged more accurately in Action (Press-Beep) than in Control (Vibration-Beep) conditions

```{r Intervals_precision}
# Precision
intErrors %>%  ggplot(., aes(y=IntervalErrorSD, x=SceneType)) +
  geom_boxplot() + facet_grid(.~EventType)
```

Interval estimates seem to be more precise (less variable) in the 3D condition.

```{r Intervals_correlation}
# Correlation, discrimination
intErrors %>%  ggplot(., aes(y=IntervalErrorCor, x=SceneType)) +
  geom_boxplot() + facet_grid(.~EventType)
```

Correlation between actual and estimated intervals seems to be a bit better in the 3D condition.

### Intervals models
- Model 1: IntervalError ~ scale(Interval) + SceneType * EventType + (1+scale(Interval) + SceneType + EventType|Subject)
- Model 2: IntervalError ~ scale(Interval) + abs(ObjectsError) + SceneType * EventType + (1+SceneType + EventType|Subject)
```{r message=FALSE, warning=FALSE}
intervalsLmer <- as.data.frame(data) %>% lmer(IntervalError ~ scale(Interval) + SceneType * EventType + (1+scale(Interval) + SceneType + EventType|Subject), .)
#summary(intervalsLmer)
#include also objects error: how inaccurate a person was in absolute numbers
intervalsOELmer <- as.data.frame(data) %>% lmer(IntervalError ~ scale(Interval) + abs(ObjectsError) + SceneType * EventType + (1+scale(Interval)+SceneType + EventType|Subject), .)
#summary(intervalsOELmer)

sjt.lmer(intervalsLmer, intervalsOELmer, file="intervalsLmer.html")# We have to save the table in html format.
htmltools::includeHTML("intervalsLmer.html") #And include it later - not needed for html_document type, only for notebook

```

- Intervals are more underestimated for longer (actual) intervals. 
- There is no effect of scene type - we hypothesise that intervals will be underestimated more in 3D than in 2D..
- The intervals may be overestimated a bit more in the Control than in Action condition - consistent with agency hypothesis, but it may be driven by factually different intervals in action than in control due to the VR SW and HW..  So we should probably rather look for an interaction:
- There is a n.s. hint at interaction between Scene type (2D/3D) and Event type (Action/Control), which is what we want - but we want there to be probably a bigger underestimation in 3D than 2D, in the Action than in the Control condition, right? 
- Model which includes also error in objects estimation (measure of difficulty of the counting task, in absolute numbers) shows no interesting difference. (The bigger significance of Interval is due to its exclusion from random effects = potentially higher false positive rate)

### Intervals RT exploration
```{r Interval_RT_exploration}
data %>% ggplot(., aes(x=IntervalLogRT, linetype=SceneType, color=EventType)) +  
  facet_wrap(~Subject) + geom_density(alpha=0.3)#geom_freqpoly(bins=5) #
```

```{r Interval_RT_Outlier_exploration, eval=FALSE, include=FALSE}
qqnorm(data$IntervalLogRT)

data.nooutliers <- data %>% group_by(Subject, SceneType, EventType) %>% 
  summarize(MeanLogRT = mean(IntervalLogRT), SDLogRT = sd(IntervalLogRT)) %>% ungroup() %>%
  mutate(Lower = MeanLogRT - 2.5*SDLogRT, Upper = MeanLogRT + 2.5*SDLogRT) %>% 
  right_join(., data, by=c("Subject", "SceneType", "EventType")) %>%
  mutate(Include = ifelse(IntervalLogRT>Lower & IntervalLogRT<Upper, T, F) ) %>% 
  filter(Include == T)

data.nooutliers %>% ggplot(., aes(x=IntervalLogRT, linetype=SceneType, color=EventType)) +  
  facet_wrap(~Subject) + geom_density(alpha=0.3)

qqnorm(data.nooutliers$IntervalLogRT)
```

RTs are around 3-7 seconds, there is a pretty long tail, but log-transformed seems ok-ish, maybe no need to remove outliers..

```{r IntervalsRT_meanlog}
intRTs <- data %>% group_by(Subject, SceneType, EventType) %>% 
  summarize(NTrials = n(), IntervalLogRTMean = mean(IntervalLogRT), IntervalLogRTSD = sd(IntervalLogRT)) 

# Mean
intRTs %>%  ggplot(., aes(y=IntervalLogRTMean, x=SceneType)) + geom_boxplot() + facet_grid(.~EventType)
```

Interval RTs seem to be faster for 2D than for 3D in Action condition

```{r IntervalsRT_sd}
# Variability
intRTs %>%  ggplot(., aes(y=IntervalLogRTSD, x=SceneType)) + geom_boxplot() + facet_grid(.~EventType)
```

Not sure whether this means something, let's wait for more data..

### Interval RT mixed-effects model
IntervalLogRT ~ Objects + Interval + SceneType * EventType + (1+SceneType + EventType|Subject)
```{r IntervalRT_mixed-effects_model, message=FALSE, warning=FALSE}
intervalRTLmer <- as.data.frame(data) %>% lmer(IntervalLogRT ~ Objects + Interval + SceneType * EventType + (1+SceneType + EventType|Subject), .)
#summary(intervalRTLmer)

sjt.lmer(intervalRTLmer, file="intervalRTLmer.html", digits.est=5, digits.ci=5)# We have to save the table in html format.
htmltools::includeHTML("intervalRTLmer.html") #And include it later, not needed for html_document type, only for notebook
```


